---
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      tidy = 'styler', 
                      comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

This RMarkdown script is designed to run chunk by chunk. You can press `ctrl+shift+Enter` to run each chunk.

Example of code chunk

```{r chunk_name}
2+2
# Run me with ctrl+shift+Enter
```
This program will download from the internet and install the latest version of the packages below if they are not already installed in your R environment. It is necessary to have internet connection to download those packages. 

If for any reason this program fails to run, please make sure that the packages are installed.

```{r R_library, include=TRUE}

#Import packages

if (!require(install.load)) {
  install.packages("install.load")
}

install.load::install_load(c("tidyverse", "readxl", "writexl", "inspectdf", "caret", "janitor", "infer", "ggchicklet"))

# remotes::install_git("https://git.rud.is/hrbrmstr/ggchicklet.git")

theme_set(theme_bw())
```

Access to the dataset can be found from the chunk below:

```{r datasets}
# Cleveland heart disease dataset

browseURL("https://archive.ics.uci.edu/ml/datasets/Heart+Disease")

# Statlog heart disease dataset
browseURL("http://archive.ics.uci.edu/ml/datasets/statlog+(heart)")

```

# Importing and preprocessing of Cleveland dataset

```{r Cleveland dataset}

# We recoded some levels of categorical features in MS-Excel for both Cleveland and Statlog heart disease database so that they can match each other

cleveland_data <- read_xlsx('data/heart_disease_database.xlsx', sheet = 1) %>% mutate(restecg = as.character(restecg))

# Data wrangling for Cleveland dataset

cleveland_data <- cleveland_data %>% 
  na_if("?") # replacing '?' to NA

# count missing value

cleveland_data %>% inspect_na() %>% show_plot()

# drop rows with NA

cleveland_data <- cleveland_data %>% drop_na() 

dim(cleveland_data)
```
  
Experiments with the Cleveland database have concentrated on simply attempting to distinguish presence (values 1,2,3,4) from absence (value 0) in the predicted attribute (num)

So we are recoding label to (1-4)- present and (0)- absent.
  
  
```{r data_processing2}
cleveland_data <- 
  cleveland_data %>% 
  mutate(heart_disease_present= factor(ifelse(num == 0, 'absent', 'present')))

# removing variable num which was the label with values 1:4

cleveland_data <- cleveland_data %>%
  select(-num)


# We changed the data type in Cleveland data to the right type 

cleveland_data <- cleveland_data %>% 
  mutate(across(!where(is_double) , as_factor))
```

# Importing and preprocessing of Statlog dataset

```{r statlog_data_data}
statlog_data <- read_xlsx('data/heart_disease_database.xlsx', sheet = 2) %>% 
  mutate(across(.cols = c(7, 12), as.character))

statlog_data <-  statlog_data %>% 
  mutate(heart_disease_present= factor(ifelse(heart_disease_present == 0, 'absent', 'present')))
                             
# Changing the data types to the right type in statlog_data data

statlog_data <-  
  statlog_data %>% 
  mutate(across(!where(is_double) , as_factor))
```

# Analysis of heart disease status across Cleveland and Statlog datasets


```{r Heart disease status analysis}
# Cleveland
cleveland_response <- cleveland_data %>%
  count(heart_disease_present) %>%
  mutate(pct = n / sum(n), data = "Cleveland") %>% rename("heart_disease_status" = "heart_disease_present") %>% 
  relocate(data)
```

```{r statlog reponse}
# Statlog
statlog_response <-  statlog_data %>%
  count(heart_disease_present) %>%
  mutate(pct = n / sum(n), data = "Statlog") %>% rename("heart_disease_status" = "heart_disease_present" ) %>% relocate(data)
```

```{r Combined heart disease database}

cleveland_statlog_response <- bind_rows(cleveland_response, statlog_response)

p <- cleveland_statlog_response %>% 
  ggplot(aes(x = heart_disease_status, y = n, fill= heart_disease_status))+ ggchicklet::geom_chicklet(show.legend = FALSE, width = 0.3)+ facet_wrap(~data)+ 
  scale_fill_manual(values = c("#1b9e77", "#d95f02"), labels = c("present", "absent"))+
  labs(x = 'Heart disease status', y = 'Number of patients') +
  geom_text(aes(label = scales::percent(pct)), vjust= -0, size= 3, colour = 'black')+
  theme(axis.title.y = element_text(size = 12, face ='bold'), axis.title.x = element_text(size = 12, face='bold'), axis.text.x = element_text(vjust = 0.3))

p

ggsave(plot = p, "Heart-disease-data-analysis/Exploratory-data-analysis/Cleveland-Statlog1.png", width = 6.74, height = 4.54)
```

# Distribution of continuous features in the Cleveland dataset by heart disease status

## rlang syntax for writing a function 

- If you have column name as a character vector (e.g. col= 'age'), use .data[[col]] i.e. column name must be with quotation  

- If the column name or expression is supplied by the user (without quotation e.g. col = age), you can pass it to aes() or vars() using {{col}} i.e. curly curly syntax.

```{r}
# Plot boxplot for continuous feature

plot_box_4contfeatures <- function(df, cols, col_x = "heart_disease_present") {
  for (col in cols) {
    p <- ggplot(df, aes(x = .data[[col_x]], y = .data[[col]], fill = .data[[col_x]])) + 
      geom_boxplot(show.legend = FALSE, width = 0.2, outlier.size = 1, outlier.shape = 5, outlier.colour = "purple") +
      scale_fill_manual(values = c("present" = "red", "absent" = "green"), aesthetics = "fill") + labs(y = str_c(col), x = NULL) + 
      theme(axis.text.x = element_text(face = "bold"), 
            axis.title.y = element_text(size = 12, face = "bold")) 
    print(p)
    
    ggsave(p, file = here::here("Heart-disease-data-analysis/Exploratory-data-analysis/box_plot/", paste0("Boxplot_", col, ".png")), width = 6.74, height = 4.54)
    
  }
}

num_cols <- 
  cleveland_data %>%
  select_if(is.numeric) %>%
  colnames()

plot_box_4contfeatures(cleveland_data, num_cols)
```

# Distribution of categorical features in the Cleveland dataset by heart disease status

```{r}
# Plot barchart for categorical features

plot_bars_4catfeatures <- function(df, cat_cols, facet_var) {
  for (col in cat_cols) {
    p <- ggplot(df, aes(x = .data[[col]], fill = .data[[col]])) +
      geom_bar(show.legend = F, width = 0.3) +
      labs(x = col, y = "Number of patients") +
      facet_wrap(vars({{ facet_var }}), scales = "free_y") +
      theme(axis.title.y = element_text(size = 12, face = "bold"), 
            axis.title.x = element_text(size = 12, face = "bold"),
            axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"))
    
    print(p)
    
    ggsave(p, file = here::here("Heart-disease-data-analysis/Exploratory-data-analysis/bar_plot/", paste0("Barplot_", col, ".png")), width = 6.74, height = 4.54)
    
  }
}

cat_cols <- 
  cleveland_data %>%
  select_if(is.factor) %>%
  colnames()

cat_cols <- cat_cols[-length(cat_cols)] # removing the class label

plot_bars_4catfeatures(cleveland_data, cat_cols, heart_disease_present)  

```


# Descriptive statistics in both Cleveland and Statlog

```{r Descriptive statistics}

# Continuous features from Cleveland heart disease data

cont_feature_ClevelandData <- cleveland_data %>% select(where(is.numeric))

cont_feature_StatlogData <-  statlog_data %>% select(where(is.numeric))


descriptive_stat <- function(tbl) {
  tbl %>% 
    vtable::st(summ = c('min(x)','max(x)', 'mean(x)','median(x)','sd(x)'), out = "return") 
}


descriptive_stat(tbl = cont_feature_ClevelandData)

descriptive_stat(tbl = cont_feature_StatlogData)
```


# Chisquare distribution table for Cleveland dataset

## One hot encoding

```{r Cleveland-Chisquare table}
# Categorical features from Cleveland heart disease data

cat_feature_ClevelandData <- cleveland_data %>%
  select(where(is.factor))

# Dummy variables 

dummies <- dummyVars("~.", data = cat_feature_ClevelandData, fullRank = TRUE)

dummy_variables <- predict(dummies, newdata = cat_feature_ClevelandData)

dummy_variables <- as_tibble(dummy_variables)

str_c(names(dummy_variables), collapse = ",")
```

The level that is not on the list of a given variable is a reference category for that variable.

## Chisquare test
```{r}
# Factor levels

fct_level <- function(fct) {
  cat_feature_ClevelandData %>% 
    select({{fct}}) %>% 
    pull %>% 
    fct_unique() %>% 
    as.character() %>% 
    str_c(collapse = ", ")
}

# Chisquare

chisquare <- function(fct) {
  cat_feature_ClevelandData %>% 
    chisq_test(response = heart_disease_present, explanatory = {{fct}}) %>% 
    pull(p_value)
}

## Feature, level, and Chi-square P-value

feature_level_tibble <-  tibble("sex" = fct_level(sex), "cp" = fct_level(cp), "fbs" = fct_level(fbs), "restecg" = fct_level(restecg),  "exang" = fct_level(exang), "slope" = fct_level(slope),  "ca" = fct_level(ca), "thal" = fct_level(thal)) %>% 
  pivot_longer(cols = everything(), names_to = "Feature", values_to = "levels")

chisq_stat_tibble <- feature_level_tibble %>% add_column(P_value = c(chisquare(sex), chisquare(cp), chisquare(fbs),chisquare(restecg), chisquare(exang), chisquare(slope), chisquare(ca), chisquare(thal))) 
```

# Contigency table showing destribution of each categorical variable by heart disease status from Cleveland data. This result was used to do Cross tabulation in `Cross_tabulation.xlsx`.

```{r}

freq_pct <- function(tbl, fct) {
  
  total <- tbl %>%  
    tabyl({{fct}}, heart_disease_present) %>% 
    adorn_totals(where = c("row", "col"))
  
  print(total)
  
  cat("\n")
  
  pct <- total %>% 
    adorn_percentages() %>% 
    adorn_pct_formatting()
  
  print(pct)
  
  }


freq_pct(cat_feature_ClevelandData, sex)

# We do this for each of the categorical features in cat_feature_ClevelandData
```


# Chisquare distribution table for Statlog dataset

## One hot encoding

```{r}

# Categorical features from Statlog heart disease data
cat_feature_StatlogData <- statlog_data %>%   select(where(is.factor)) 
  
# Dummy variables 

dummy_variables <- predict(dummies, newdata = cat_feature_StatlogData)

dummy_variables <- as_tibble(dummy_variables)

str_c(names(dummy_variables), collapse = ",")
```

The level that is not on the list of a given variable is a reference category for that variable.

## Chisquare test

```{r}
# Factor levels

fct_level <- function(fct) {
  cat_feature_StatlogData %>% 
    select({{fct}}) %>% 
    pull %>% 
    fct_unique() %>% 
    as.character() %>% 
    str_c(collapse = ", ")
}


# Chisquare

chisquare <- function(fct) {
  cat_feature_StatlogData %>% 
    chisq_test(response = heart_disease_present, explanatory = {{fct}}) %>% 
    pull(p_value)
}

## Feature, level, and Chi-square P-value

feature_level_tibble <-  tibble("sex" = fct_level(sex), "cp" = fct_level(cp), "fbs" = fct_level(fbs), "restecg" = fct_level(restecg),  "exang" = fct_level(exang), "slope" = fct_level(slope),  "ca" = fct_level(ca), "thal" = fct_level(thal)) %>% 
  pivot_longer(cols = everything(), names_to = "Feature", values_to = "levels")

chisq_stat_tibble <- feature_level_tibble %>% add_column(P_value = c(chisquare(sex), chisquare(cp), chisquare(fbs),chisquare(restecg), chisquare(exang), chisquare(slope), chisquare(ca), chisquare(thal))) 
```

# Contigency table showing distribution of each categorical variable by heart disease status from Statlog data


```{r}
freq_pct(cat_feature_StatlogData, cp)

# We do this for each of the categorical features in cat_feature_StatlogData

```

# The end!